### 2011-02-07 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
first version of the deliverable for M36
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1252 d8b0c59b-6544-0410-94d9-83a5658a058e

\usepackage{inputenc}
\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\parindent0pt

\newcommand\deliverableNumber{D5.1.2}
\newcommand\deliverableTitle{User Manual of the FAST Catalogue}
\newcommand\deliverableTitleShort{Catalogue User Manual}
\newcommand\workpackageNumber{5}
\newcommand\workpackageTitle{Semantic catalogue of screen-flow resources and back-end Web Services}
\newcommand\authorOne{Ismael Rivera, NUIG}
\newcommand\authorTwo{Author 2}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

@@

\pagenumbering{roman}

@@

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 25/01/2010}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
\rofoot{\small{\thepage}}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                         \textcolor{white}{\textbf{Date}} &
                         \textcolor{white}{\textbf{Author (Partner)}} &
                         \textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 25/01/2010 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
\end{tabular}
\end{small}

@@

\begin{small}
\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable is the technical documentation for the prototype developed as part of the WP5. It describes the catalogue's architecture, data interchange formats, APIs and examples of how to interact with it.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\
  \hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{NUIG} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495338} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, semantic catalogue, gadget catalogue, RDF store}\\ \hline
\end{tabular}
\end{small}

@@

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{2}
\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGINNING OF SECTIONS %
%%%%%%%%%%%%%%%%%%%%%%%%%
\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}

@@

The deliverable presents both the external and internal architecture in Section~\ref{sec:architecture_overview}, then in Section~\ref{sec:restful_catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.

@@

This should be the first version of this deliverable since in M12 the deliverable 5.1 was not scheduled. Nevertheless, last year an extra document was written as an appendix to the FAST Catalogue in order to detail and clarify technical decisions and to have a technical reference manual. Hence, the deliverable D5.2.1 will evolve to become the D5.1.2, so it will be no longer needed by itself.

There are several new sections and major changes which need to be mentioned:

\begin{itemize}
	\item Section~\ref{sec:installation_guide}. Installation Guide describes a set of steps to be followed in order to install and execute the FAST Catalogue. Also some configuration instructions are specified.
	\item Section~\ref{sub:content_negotiation}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
	\item Section~\ref{sec:restful_catalogue_api}. RESTful Catalogue API has been intensively extended and modified in order to  reflect all the new functionality and how every request or response is constructed. Some of the most important features added in this iteration are:
	\begin{itemize}
		\item support for form elements, operators and back-end services (see Section ...),
		\item convenience operations for recommendation and checking satisfaction and reachability for form elements, operators and back-end services (see Section ...),
		\item and, last but not least, support for Screen planning (see Section~\ref{sub:planning}).
	\end{itemize}
	\item Several appendixes have been added with the JSON structure of every building block the Catalogue can deal with.
	\item Appendix A. Last year an evaluation to choose the best technology for the Catalogue was done. It was not included in any formal deliverable, so this year it has been documented as an appendix.
\end{itemize}

@@

Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed. If you are testing the application in your machine, it may be something like \url{http://localhost:8080/FASTCatalogue}.

@@

\begin{verbatim}
storageDir=D:\\FAST\\catalogue\\repository
\end{verbatim}

@@

\begin{verbatim}
sesameServer=http://localhost:8880/openrdf-sesame
repositoryID=c2
\end{verbatim}

@@

It is recommended to check if the FAST Catalogue has been successfully installed. To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of resource such as Screen. The URL would be similar to \url{http://localhost:8080/FASTCatalogue/screens}. Depending on the variant chosen, the response may be different. For JSON, the response should be \verb|[]|, but in HTML the message \verb|No screens found.| should be shown.

@@

This section presents a technical overview of the internal catalogue's architecture and which external components will interact with it. Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure~\ref{fig:catalogue_architecture}).

@@

The Presentation layer will be the public interface of the catalogue. The main purpose of this layer is to provide its functionality to other components within the Gadget Visual Studio. The service is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs. REST is an architectural style, not a toolkit or a standard. Even though, it makes use of standards like HTTP, URI, Resource Representations (XML, HTML, JSON, JPEG, etc.) or MIME types (text/xml, text/html, application/json, image/jpeg, etc.). Another characteristic of the REST style is its stateless assumption. The catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response.

As part of the Presentation layer, a SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification. The SPARQL endpoint is mostly offered to enable other developers to query directly the catalogue knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.

The Business Logic layer contains all the FAST domain-specific processing. It provides functions to interact with all the elements of the domain model specified in \cite{moeller2009fast_ontology}. It processes the input from the Presentation layer, creating specific objects modelling the business logic, and vice versa. In addition it interacts with the Persistence layer in order to persist the model.

The Persistence layer provides an API which allows the Business layer to work with a standard set of objects that read and save their state to the triple store. For this reason, on top of this layer it is used RDF2Go, an abstraction over triple (and quad) stores. The RDF2Go API allows interacting with the semantic representation of the model (triples) in a generic manner, and also brings the flexibility of choosing different triple stores to persist them. The selected implementation for the triple stored is Sesame 2. Moreover, this framework is completely extensible and configurable regarding to storage mechanisms, inferencers, RDF file formats, query result formats and query languages. For the actual catalogue prototype, the storage mechanism used is the native file storage system of Sesame, the inferencer used is a subset of the RDFS entailment rules \cite{rdfsrules} following a forward-chaining policy.

@@

\clearpage
\section{RESTful Catalogue API} % (fold)
\label{sec:restful_catalogue_api}

@@

\singlespacing
\begin{verbatim}
{
  "canvas": [
    { "uri": "http://localhost:8080/catalogue/screens/238" },
    { "uri": "http://localhost:8080/catalogue/screens/323" }
  ],
  "elements": [
    { "uri": "http://localhost:8080/catalogue/screens/636" }
  ],
  "domainContext": {
    "tags": [
      {
        "label": { "en-GB": "Amazon" },
        "means": "http://dbpedia.org/page/Amazon.com"
      }
    ],
    "user": "irivera",
  },
  "criterion": "reachability"
}
\end{verbatim}
\doublespacing

@@

\singlespacing
\begin{verbatim}
{
  "label": {
    "en-GB": "Simple Example",
    "es-ES": "Ejemplo Sencillo",
    "de-DE": "Einfaches Beispiel"
  },
  "description": {
    "en-GB": "This is a simple example of a screen",
    "es-ES": "Esto es un ejemplo sencillo de pantalla",
    "de-DE": "Dies ist ein einfaches Beispiel für einen Bildschirm"
  },
}
\end{verbatim}
\doublespacing

\singlespacing
\begin{verbatim}
<rdf:Description rdf:about="http://www.morfeoproject.eu/fast/fco#ExampleScreen">
  <rdf:type rdf:resource="http://www.morfeoproject.eu/fast/fco#Screen"/>
  <rdfs:label xml:lang="en-GB">Simple Example</rdfs:label>
  <rdfs:label xml:lang="es-ES">Ejemplo Sencillo</rdfs:label>
  <rdfs:label xml:lang="de-DE">Einfaches Beispiel</rdfs:label>
  <description
    xmlns="http://purl.org/dc/terms/"
    xml:lang="en-GB">This is a simple example of a screen</description>
  <description
    xmlns="http://purl.org/dc/terms/"
    xml:lang="es-ES">Esto es un ejemplo sencillo de pantalla</description>
  <description
    xmlns="http://purl.org/dc/terms/"
    xml:lang="de-DE">Dies ist ein einfaches Beispiel für einen Bildschirm</description>
</rdf:Description>
\end{verbatim}
\doublespacing

Until now, using internationalisation is allowed for the label and the description of a screen, but in the near future, more attributes will benefit from this feature.

@@

\subsection{Content Negotiation} % (fold)
\label{sub:content_negotiation}

Before any major details about how content negotiation is addressed by the catalogue, a few concepts need to be understood. First of all, say a URI is not a file name. It is common to see a URI as the location for a file, but a URI or Universal Resource Identifier is not a URL: it is not a file name or location, it is an identifier (or a reference) to a resource. Then by dereferencing a URI the actual resource can be retrieved. Let's see a request to retrieve the resource "Screen 123" is sent to the server. The resource is just a piece of information, however the server will answer an HTML document with a text describing the screen or an image representing it. On the Web, these equivalent representations of a resource are called variants, and content negotiation is the mechanism used to determine which of the representations of most appropiate for a given request.

@@

In summary, the basic idea of content negotiation, as stated in the HTTP 1.1 specification ~\cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.

In fact, what the catalogue really does is "format negotiation" since the alternate representationos are just based on the selection of the media type, through the Accept header, but does not consider different languages or encoding types. The formats supported are JSON, RDF+XML and HTML.

@@

This section contains detailed descriptions of the interface provided by the catalogue via REST services, detailing request parameters, response elements, any special errors and examples of requests and responses. The URL format is also specified for each operation, where 'catalogueURL' has to be replaced for the real URL the service is installed (e.g. http://demo.fast.morfeo-project.org/catalogue).

\subsubsection{CRUD operations} % (fold)

Any building block within the Catalogue is a resource in terms of REST philosophy, hence it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation. Two concepts have to be defined: \emph{collection}, as a set of resources which access URI is http://catalogueURL/<type>/ where catalogueURL is the specific URL where the Catalogue server is installed and <type> is the plural of the name of the building block (e.g. screens, services), and \emph{member} of the collection, in other words, the building block itself, which access URI is http://catalogueURL/<type>/<id> where the <id> has to be replaced by the identifier of a specific building block. The details of the operations and which HTTP verb has to be used can be found in the Table~\ref{tab:crud_operations}.

@@

To create a new screen, a POST request is send to the catalogue server, including the JSON representation of the screen as the body of the request:

>>>>

\usepackage{inputenc}
\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\usepackage{amsmath}

% This helps to get hyphenation in typewriter font (\texttt)
\usepackage[htt]{hyphenat}

\parindent0pt

\newcommand\deliverableNumber{D5.1.2}
\newcommand\deliverableTitle{User Manual of the FAST Catalogue}
\newcommand\deliverableTitleShort{Catalogue User Manual}
\newcommand\workpackageNumber{5}
\newcommand\workpackageTitle{Semantic catalogue of screen-flow resources and back-end Web Services}
\newcommand\authorOne{Ciprian Palaghita, Cyntelix}
\newcommand\authorTwo{Ismael Rivera, NUIG}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

@@

% \pagenumbering{roman}

@@

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		\vspace{0.2cm}
		{\large \sffamily \authorTwo}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 17/02/2011}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
% \cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
% \rofoot{\small{\thepage}}

\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort{} $\bullet$ Page \thepage\ of \pageref{LastPage}}}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{6.6cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                         \textcolor{white}{\textbf{Date}} &
                         \textcolor{white}{\textbf{Author (Partner)}} &
                         \textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 19/02/2010 & Ciprian Palaghita (Cyntelix) & First version of the deliverable \\
 & & Ismael Rivera (NUIG) & \\ \hline
2.0 & 17/02/2011 & Ciprian Palaghita (Cyntelix) & Second version of the deliverable \\
 & & Ismael Rivera (NUIG) & \\ \hline
\end{tabular}
\end{small}

@@

\begin{small}
\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.2cm}|p{1.6cm}|p{3.37cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {25/02/2011}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable is the technical documentation for the prototype developed as part of the WP5. It describes the catalogue's architecture, data interchange formats, APIs and examples of how to interact with it.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
	{} & \multicolumn{4}{l|}{\authorTwo}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\
  \hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{Cyntelix} & \textbf{email} & {cpalaghita@cyntelix.com} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 858494258} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, semantic catalogue, gadget catalogue, RDF store}\\ \hline
\end{tabular}
\end{small}

@@

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{2}
\tableofcontents
\cleardoublepage
% \pagenumbering{arabic}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGINNING OF SECTIONS %
%%%%%%%%%%%%%%%%%%%%%%%%%
% \rofoot{\small{Page \thepage\ of \pageref{LastPage}}}

@@

The deliverable presents both the external and internal architecture in Section~\ref{sec:architecture_overview}, then in Section~\ref{sec:catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.

@@

changes from M24 version:
\begin{itemize}
 \item 'find' and 'check' operations for screens (screen-flow design) has been removed. the same functionaly was offered by 'find\&check' doing just 1 request.
 \item API calls (both find\&check, planner, metadata) updated to reflect the changes in the prototype
 \item Adds operation for building block cloning.
 \item added appendix explaining how the recommendation/planning is done
 \item architecture have been modified reflecting the changes in the catalogue's architecture, and including new components such as the machine learning library (Mahout) and a library to query iServe.
 \item uses Sindice in the 'ontology importer' mechanism
\end{itemize}

@@

Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed. If you are testing the application in your machine, it may be something like \verb|http://catalogueURL/FASTCatalogue|.

@@

\begin{verbatim}
storageDir=/opt/fast-catalogue/repository
\end{verbatim}

@@

\begin{verbatim}
sesameServer=http://localhost:8880/openrdf-sesame
repositoryID=repository-ex1
\end{verbatim}

@@

It is recommended to check if the FAST Catalogue has been successfully installed. To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of building block such as Screen. The URL would be similar to \verb|http://catalogueURL/FASTCatalogue/screens|. Depending on the variant chosen, the response may be different. For JSON, the response should be \verb|[]|, but in HTML the message \verb|No screens found.| should be shown.

@@

This sections presents a technical overview of the Catalogue's architecture (see Figure~\ref{fig:catalogue_architecture}), and how the interaction with external applications or APIs is done. The Catalogue platform has a modular design, which are spread in basically three layers, depending on the nature of each component:
\begin{itemize}
 \item The top layer represents the public API for the Catalogue. It puts together all the available information and services to offer the maximum powerful and flexibility for other IDEs or applications. In FAST, here is where the Gadget Visual Storyboard
 \item The core layer implements the object model, and provides the business logic for the different processes involved in the Catalogue, integrating third party libraries where needed.
 \item The pumbling layer offers an abstraction to RDF model logic and triple stores.
\end{itemize}

@@

\subsection{Presentation layer / API}

The Presentation layer will be the public interface or API of the Catalogue. The main purpose of this layer is to provide its functionality to the Gadget Visual Storyboard, or any other third party IDE or application. The API is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs. REST is an architectural style, not a toolkit or a standard. Even though, it makes use of standards like HTTP, URI, Resource Representations or MIME types. Another characteristic of the REST style is its stateless assumption. The Catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response. The API provided is explained in detail in Section~\ref{sec:catalogue_api}.

As part of the Presentation layer, a SPARQL endpoint has been implemented using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification. The SPARQL endpoint is mostly offered to enable other developers to query directly the Catalogue knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.

\subsection{Core layer}

The Core or Business Logic layer contains all the FAST domain-specific processing. The resources and relations between each other, which define the domain model, are extracted from the FAST ontology~\cite{moeller2011fast_ontology}. These layer provides support to the functions defined in the API, and interact with the persistence layer to retrieve and store any information of the data model.

This layer provides several mechanisms to search and browse these different resources. To know more about these capabilities, please read Sections~\ref{ssub:screen_findcheck},~\ref{ssub:screen_component_findcheck} and~\ref{sub:planning} where the API calls are defined, and Appendix F for more details about the implementation.

At the bottom part of the Gadget Architecture~\cite{reyes2011gadget_architecture} are the back-end services. These are RESTful services, SOAP-based web services, or similar. In FAST, these web services are wrapped, extracting their definition and encapsulation their behaviour. To creation of these service wrappers is on demand, and the input for this process is a particular definition of a web service. However, the Catalogue does not crawl the Web looking for Web services. Instead, and to enrich the collection of available web services, the Catalogue integrates and queries iServe\footnote{http://iserve.kmi.open.ac.uk/}, a platform for publishing Semantic Web Services as linked data, no matter their original format. Therefore, thousands of web services defined or annotated in a semantic way, are also served from the Catalogue, in a uniform manner.

The Catalogue is a semantic application. As a result, it has to handle with ontologies. Apart from the core ontologies being used across the whole application (see~\cite{moeller2011fast_ontology}), the building blocks may reference classes and properties from diverse vocabularies and ontologies. To have a better understanding of them, the Catalogue imports them inside its repository any time it detects a building block is using an unknown class, property, or in other words, an unknown ontology. While importing an ontology, the Catalogue tries to get its definition from the URI of a given class or property (as stated in~\cite{berrueta2008}). However, sometimes this may be not succesful, and the Catalogue will try to get its definition querying Sindice\footnote{http://www.sindice.com/}, a semantic search engine, which cralws the web and keeps provenance metadata for the ontologies.

\subsection{Plumbing layer}

The Plumbing layer abstract the interaction with triple stores, RDF models and physical persistence to disk, database, etc. It provides the different mechanisms to allow the Catalogue the storage of the domain objects into a triple store in its underlying representation of RDF.

There are many triple stores solutions, any of them with its own advantages and disadvantages. To make the Catalogue independent of the solution adopted, it makes uses of RDF2Go, an abstraction over triple (and quad) stores. The RDF2Go API allows interacting with the semantic representation of the model (triples) in a generic manner, and also brings the flexibility of choosing different triple stores to persist them.

The current implementation of the Catalogue prototype at M36 uses Sesame 2 as triple store. This solution has a wide community behind, and it has attractive advantages such as being completely extensible and configurable regarding to storage mechanisms, inferencers, RDF file formats, query result formats and query languages.

@@

\clearpage
\section{Catalogue API} % (fold)
\label{sec:catalogue_api}

@@

\singlespacing
\begin{verbatim}
{
  "canvas": [
    { "uri": "http://catalogueURL/screens/238" },
    { "uri": "http://catalogueURL/screens/323" }
  ],
  "palette": [
    { "uri": "http://catalogueURL/screens/636" }
  ],
  "domainContext": {
    "tags": [
      {
        "label": { "en-GB": "Amazon" },
        "means": "http://dbpedia.org/page/Amazon.com"
      }
    ],
    "user": "http://ismaelrivera.es/#me",
  },
  "criterion": "reachability"
}
\end{verbatim}
\doublespacing

@@

\singlespacing
\begin{verbatim}
{
  "label": {
    "en-GB": "Product Search",
    "es-ES": "Búsqueda de productos",
    "de-DE": "Produktsuche"
  },
  "description": {
    "en-GB": "On this screen, the user can enter a search term.",
    "es-ES": "En esta ventana, el usuario puede introducir el criterio de búsqueda",
    "de-DE": "Auf diesem Screen kann der Benutzer einen Suchbegriff eingeben."
  }
}
\end{verbatim}
\doublespacing

\singlespacing
\begin{verbatim}
<rdf:RDF
    xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
    xmlns:foaf="http://xmlns.com/foaf/0.1/"
    xmlns:dc="http://purl.org/dc/terms/"
    xmlns:fgo="http://purl.oclc.org/fast/ontology/gadget#"
    xmlns:xsd="http://www.w3.org/2001/XMLSchema#"
    xmlns:rdfs="http://www.w3.org/2000/01/rdf-schema#"
    xmlns:ctag="http://commontag.org/ns#">
  <rdf:Description rdf:about="http://demo.fast.morfeo-project.org/FASTCatalogue/screens/1">
    <rdf:type rdf:resource="http://purl.oclc.org/fast/ontology/gadget#Screen"/>
    <rdfs:label xml:lang="en-gb">Product Search</rdfs:label>
    <rdfs:label xml:lang="en-es">Búsqueda de productos</rdfs:label>
    <rdfs:label xml:lang="en-de">Produktsuche</rdfs:label>
    <dc:description xml:lang="en-gb">On this screen, the user can enter a search
    term.</dc:description>
    <dc:description xml:lang="en-gb">En esta ventana, el usuario puede introducir
    el criterio de búsqueda.</dc:description>
    <dc:description xml:lang="en-gb">Auf diesem Screen kann der Benutzer einen
    Suchbegriff eingeben.</dc:description>
  </rdf:Description>
</rdf:RDF>
\end{verbatim}
\doublespacing

Internationalisation is being used by the attributes \emph{label} and \emph{description} of any building block, \emph{label} of the pre-/post-conditions and \emph{label} of the tags.

@@

\subsection{Linked Data} % (fold)
\label{sub:linked_data}

Apart from serving as the back-end system for the Gadget Visual Storyboard (GVS), the catalogue publishes all FAST building blocks as linked data, following the principles as originally defined in \cite{bernersLee2006linkedData}, in order to make them available to arbitrary third party applications. Each building block is identified by an HTTP URI and hosted through the catalogue so that it can be dereferenced through the same URI. For each building block, data is available in representations in different standard formats such as JSON (for communication with the GVS), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version. 
Following the principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris}, these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}.
As required by the forth rule of linked data, individual building blocks link to other data on the Web, thereby preventing so-called isolated ``data islands''.

In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood. The URI of a particular building block must be understood as the identifier of the building block as such, as opposed to a particular representation (JSON, HTML, etc.) of it. In fact, each such representation has its own URI, under which it can be retrieved. Using the terminology established in \cite{w3c2004webArchitecture}, the building block is a \emph{non-information resource}, whereas each representation is an \emph{information resource}. In the process of content negotiation, the requesting agent will first request the URI of the building block as such (e.g., \verb|http://catalogueURL/screens/235|), as well as the required representation format (e.g., JSON). The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \verb|http://catalogueURL/screens/235.json|), as illustrated in Fig.~\ref{fig:content_negotiation}.
On the Web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.

@@

In summary, the basic idea of content negotiation, as stated in the \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences. In HTTP, this is done by the client which may send, in its request, accept headers (\texttt{Accept}, \texttt{Accept-Language} and \texttt{Accept-Encoding}), to communicate its capabilities and preferences in format, language and encoding, respectively.

In fact, what the catalogue really does is ``format negotiation'' since the alternate representations are just based on the selection of the media type, through the accept header, but does not consider different languages or encoding types. The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa. Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI. The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.json} or \texttt{.html}.

@@

The Catalogue API is a programmatic interface into many of the Catalogue's features. It includes not only control over creation, modification and deletion of building blocks, but it has been extended to support advanced search and recommendation capabilities.

This section contains detailed descriptions of the operations permitted by the API, detailing request parameters, response elements, any special errors and examples of requests and responses. The URL format is also specified for each operation, where \emph{catalogueURL} has to be replaced for the real URL the service is installed (e.g. http://demo.fast.morfeo-project.org/catalogue).

\subsection{Managing Building Blocks} % (fold)
\label{sub:managing_building_blocks}

Any building block within the Catalogue is a resource in terms of REST philosophy, therefore it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation. Two concepts should be clarified: \emph{collection}, as a set of resources which access URI is \verb|http://catalogueURL/[type]| where \emph{catalogueURL} is  where the Catalogue server is installed and \verb|[type]| is the plural of the name of the building block (e.g. screens, operators), and \emph{member} of the collection, in other words, the building block itself, which access URI is \verb|http://catalogueURL/[type]/[id]| where \verb|[id]| has to be replaced by the identifier of a specific building block. The details of the operations and which HTTP verb has to be used can be found in the Table~\ref{tab:crud_operations}. These methods are to what is called \emph{Prototype} of building blocks (see \cite{moeller2011fast_ontology}).

@@

To create a new screen, a POST request is sent to the catalogue server, concretely to \texttt{http://cata\-logueURL/screens}, including the JSON representation of the screen in the body of the request:

\singlespacing
\begin{verbatim}
{
  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
  "creationDate": "2010-01-26T17:01:13+0000",
  "creator": "http://ismaelrivera.es/#me",
  "description": {"en-gb": "Please fill the description..."},
  "homepage": "http://fast.morfeo-project.eu/",
  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
  "id": "28",
  "label": {"en-gb": "Amazon Product List"},
  "postconditions": [
    {
      "id": "item",
      "label": { "en-gb": "An item" },
      "pattern": "?item 
                  http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                  http://fast.morfeo-project.org/ontologies/amazon#Item",
      "positive": "true"
    }
  ],
  "preconditions": [
    {
      "id": "filter",
      "label": { "en-gb": "A search request" },
      "pattern": "?search
                  http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                  http://fast.morfeo-project.org/ontologies/amazon#SearchRequest",
      "positive": "true"
    }
  ],
  "rights": "http://creativecommons.org/",
  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
  "tags": [{"label": {"en-gb": "amazon"}}],
  "version": "1.0"
}
\end{verbatim}
\doublespacing
